### 2023-12-17 thesis/chapters/02-analysis-A-CdM.tex
Write data exploration in design

Measurement of vibrations should be accompanied by a description of the machine and its operating conditions. The machine description includes its identifier and type, power source, rated rotation speed and power, configuration (shaft or belt driven), and machine support. Measurement parameters are to be recorded alongside the measurement value itself such as timestamp, transducer type, sensor location and orientation in MIMOSA code, measurement units and units qualifier (p-p, RMS), and other processing options (filters, number of averages, etc.)~\cite{noauthor_iso_2002}.

>>>>

Measurement of vibrations should be accompanied by a description of the machine and its operating conditions. The machine description includes its identifier and type, power source, rated rotation speed and power, configuration (shaft or belt driven), and machine support. Measurement parameters are to be recorded alongside the measurement value itself such as timestamp, transducer type, sensor location and orientation in \textbf{MIMOSA convention} (Machinery Information Mangement Open System Alliance), measurement units and units qualifier (p-p, RMS), and other processing options (filters, number of averages, etc.)~\cite{noauthor_iso_2002}.

### 2023-12-17 thesis/chapters/02-analysis-D-datasets.tex
Write data exploration in design

Observations were conducted in three cardinal axes simultaneously with 2 sets of accelerometers each one associated with one bearing (inner and outer bearings)~(Fig.~\ref{fig:mafaulda-simulator}). Additionally, a magnetic tachometer produced a pulse on shaft turn. The cardioid condenser microphone recorded sound emissions with a frequency range 20~Hz - 20~kHz. Sensors were fed into a four-channel dynamic signal acquisition module.

@@

\begin{table}[h]
\renewcommand{\arraystretch}{1.2}
\centering
\begin{tabular}{|l|l|}
\hline
\textbf{Columns} & \textbf{Description}                                                                                                                                               \\ \hline
1.               & \begin{tabular}[c]{@{}l@{}}Pulse with modulation of tachometer signal \\ to estimate rotation frequency  (in TTL levels)\end{tabular}                              \\ \hline
2., 3., 4.       & \begin{tabular}[c]{@{}l@{}}Underhang bearing accelerometer \\ (inner - between the rotor and motor)\\ - axial, radial, tangential direction\end{tabular}           \\ \hline
5., 6., 7.       & \begin{tabular}[c]{@{}l@{}}Overhang bearing accelerometer \\ (outer - outside most position after the rotor)\\ - axial, radial, tangential direction\end{tabular} \\ \hline
8.               & Microphone                                                                                                                                                         \\ \hline
\end{tabular}
\caption{MaFaulDa description of columns}
\label{tab:mafaulda-columns}
\end{table}

@@

The accelerometers used are piezoelectric and have a frequency range of up to 10 KHz, dynamic range of $\pm$490 $m/s^2$, and resolution step of 10.2~mV per $m/s^2$. These sensor parameters are the same as in the case of MaFaulDa. In total, three different uniaxial accelerometers are mounted on the motor housing.

>>>>

Observations were conducted in three cardinal axes simultaneously with 2 sets of accelerometers each one associated with one bearing (inner and outer bearings)~(Fig.~\ref{fig:mafaulda-simulator}). Additionally, a magnetic speedometer produced a pulse on shaft turn. The cardioid condenser microphone recorded sound emissions with a frequency range 20~Hz - 20~kHz. Sensors were fed into a four-channel dynamic signal acquisition module.

@@

\begin{table}[h]
\renewcommand{\arraystretch}{1.2}
\centering
\begin{tabular}{|l|l|}
\hline
\textbf{Columns} & \textbf{Description}                                                                                                                                               \\ \hline
1.               & \begin{tabular}[c]{@{}l@{}}Pulse with modulation of speedometer signal \\ to estimate rotation frequency  (in TTL levels)\end{tabular}                              \\ \hline
2., 3., 4.       & \begin{tabular}[c]{@{}l@{}}Underhang bearing accelerometer \\ (inner - between the rotor and motor)\\ - axial, radial, tangential direction\end{tabular}           \\ \hline
5., 6., 7.       & \begin{tabular}[c]{@{}l@{}}Overhang bearing accelerometer \\ (outer - outside most position after the rotor)\\ - axial, radial, tangential direction\end{tabular} \\ \hline
8.               & Microphone                                                                                                                                                         \\ \hline
\end{tabular}
\caption{MaFaulDa description of columns}
\label{tab:mafaulda-columns}
\end{table}

@@

The accelerometers used are piezoelectric and have a frequency range of up to 10 kHz, dynamic range of $\pm$490 $m/s^2$, and resolution step of 10.2~mV per $m/s^2$. These sensor parameters are the same as in the case of MaFaulDa. In total, three different uniaxial accelerometers are mounted on the motor housing.

### 2023-12-17 thesis/chapters/03-design.tex
Write data exploration in design

\noindent In accomplishing the objectives of our research we propose several \textbf{goals}:
\begin{itemize}
    \itemsep0pt
	\item Statistically and visually describe vibration signals from the Machinery fault database (MauFaulDa).
	\item Establish a list of conditions that should be later investigated in the experimental setting.
	\item Prepare dataset to be used in conjunction with statistical learning models, namely by identifying labels and balancing classes.
	\item Find the best subsets of features in temporal and spectral domain with previously analyzed feature extraction and selection methods.
	\item Evaluate the performance of models described in the diagnostics section with a significant focus on the k-nearest neighbor algorithm.
	\item Combine feature selection with online machine learning model.
	\item Acquire measurements of vibrations from machines in the real environment to form a novel dataset of machinery behavior.
	\item Develop hardware and implement its firmware to obtain such measurement in the quality demanded by vibrodiagnostics standards.
\end{itemize}

However, we leave out from our efforts experiments on the data features from wavelets and on the peaks in the spectral domain. The reason is that we did not find a way to represent extracted features more succinctly as a one number. We also did not discover a strategy for choosing only the relevant frequency bins. The assembled description is kept to empower further research on that topic.

The tasks are associated with \textbf{risks} impacting their successful completion. The possible risks are assessed and prioritized. The most notable risk encountered early on is that the machine in the real environment will not be available for vibration measurements. We successfully eliminated the risk by contacting and establishing collaboration with alternative partners. The additional risks are that the repeated measurements will not be consistent and fault modes could not be reliably differentiated and labeled in the dataset because each machine is unique in its structure. Related is the risk that not enough data is obtained from various classes.  Also, suggestions can made by exploring the MaFaulDa dataset which will not be applicable in the practice. All these risks have to be regularly reevaluated to achieve our goals.

\section{Dataset exploration}

\begin{figure}[ht]
	\centering
	\includegraphics[width=\textwidth]{assets/design/activity-data-exploration.png}
	\caption{Activity diagram of MaFaulDa dataset preprocessing}
\end{figure}

\begin{figure}[ht]
    \centering
    \begin{subfigure}[b]{0.44\textwidth}
        \includegraphics[width=\textwidth]{assets/design/Mafaulda-A-time-waveform.png}
        \caption{Temporal domain waveforms in all spatial directions with 300 ms duration}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.55\textwidth}
        \includegraphics[width=\textwidth]{assets/design/Mafaulda-A-spectrum-Y-axis.png}
        \caption{Frequency spectrum in radial direction (y)}
    \end{subfigure} 
    \caption{Signals at 1 s timestamp from inner bearing in each fault category with the highest fault severity}
\end{figure}

>>>>

\noindent In accomplishing the objectives of our research we propose several \textbf{goals}:
\begin{todolist}
    \itemsep0pt
	\item Statistically and visually describe vibration signals from the Machinery fault database (MauFaulDa).
	\item Establish a list of conditions that should be later investigated in the experimental setting.
	\item Prepare dataset to be used in conjunction with statistical learning models, namely by identifying labels and balancing classes.
	\item Find the best subsets of features in temporal and spectral domain with previously analyzed feature extraction and selection methods.
	\item Evaluate the performance of models described in the diagnostics section with a significant focus on the k-nearest neighbor algorithm.
	\item Combine feature selection with online machine learning model.
	\item Acquire measurements of vibrations from machines in the real environment to form a novel dataset of machinery behavior.
	\item Develop hardware and implement its firmware to obtain such measurement in the quality demanded by vibrodiagnostics standards.
\end{todolist}

However, we leave out from our efforts experiments on the data features calculated from wavelets and peaks in the spectral domain. The reason is that we did not find a way to represent extracted features more succinctly as a single number. We also did not discover a strategy for choosing only the relevant frequency bins. The assembled description is retained to lead further research on that topic. \\

The tasks are associated with certain \textbf{risks} impacting their successful completion. The possible risks are assessed and prioritized. The most notable risk encountered is that the machine in the real environment will not be available for vibration measurements. We successfully eliminated the risk by contacting and establishing collaboration with alternative partners.

The additional risks are that the repeated measurements will not be consistent and fault modes could not be reliably differentiated and labeled in the dataset because each machine is unique in its structure. Related is the risk that not enough data is obtained from various classes and suggestions made by exploring the MaFaulDa dataset not be applicable in practice. All these risks have to be tracked and regularly reevaluated to achieve our goals.

\section{Dataset exploration}
In establishing the validity of methods to be deployed on the sensor node we explore the MaFaulDa dataset. It is the largest known machinery fault collection, so it is possible to create multiple subsets based on requested conditions.

One representative recording is first selected in each available fault category. The sample is visualized and statistically described in both temporal and spectral domains. A whole step-by-step procedure is outlined in the activity diagram (Fig.~\ref{fig:design:mafaulda-preprocessing}).

\begin{figure}[ht]
	\centering
	\includegraphics[width=\textwidth]{assets/design/activity-data-exploration.png}
	\caption{Activity diagram of MaFaulDa dataset preprocessing}
	\label{fig:design:mafaulda-preprocessing}
\end{figure}

MaFaulda contains 1951 records labeled with inducted faults of increasing severity. The defects were set up on the machine simulator as is mentioned in the part about datasets. Time series of the triaxial piezoelectric accelerometers in separate files have a sampling frequency of 50 kHz.

These vibration sensors are placed in two positions. The first placement is around the inner underhang bearing named \emph{A} which is closer to the motor. The second location is around the outer overhang bearing denoted as \emph{B} position.

\subsection{Fault annotations}
The dataset has annotations altogether for 10 classes of faults of which there is 1 class for fault-free baseline operation, 3 classes for shaft defects, 3 classes for inner bearing defects, and 3 for outer bearing defects. Some categories are redundant or irrelevant for a given sensor position.

Therefore rotor shaft misalignements in vertical and horizontal directions are merged into one joint group. Depending on the chosen bearing position only records having relevant labels are considered.

This means that fault classification solely concerns bearing in the direct contact and shaft mechanically passing through it. The bearings affect each other, but the effect on the opposite side should appear via a common interconnection shaft.

In the end, that leaves \textbf{6 types of labels}: baseline, two shaft faults are imbalance and misalignment, and three bearing faults are cage fault, ball fault, and outer race fault. In the step of choosing the accelerometer location records pointing to the other bearing as a source of the malfunction are discarded. Then the next action renames the labels to be better recognizable and unite the same phenomena.

Groups of identified machine defects are additionally characterized by altered masses attached or motor shaft displacement shifts. The set amount is sorted in ascending order separating \textbf{multiple event severities}. However, the count of severity levels is not identical in every group. Levels are hence scaled into the range between zero and one using a min-max scaler. Scaling is applied to classes separately.

The strength of the recorded response by the underlying defect is also dependent on the shaft \textbf{rotational speed}. Speed in rpm is calculated from pulsed speedometer output. It is the average distance between two successive rising edges: 
\begin{ceqn}\begin{align}
\mathrm{rpm} = 60 \;/\; \overline{\Delta t}
\end{align}\end{ceqn}

\textbf{A specimen waveform} is picked from each fault class to illustrate their superficial differences. Recordings are filtered to get one of the highest severity levels and around a mean rotation speed of 2500 rpm (42 Hz)  to see the patterns most pronounced. The baseline class sample is chosen according to fixed rotor speed.

\subsection{Signal filters}
The DC component in the three-dimensional vibration signal is removed by subtracting the global mean. Immediately follows a digital IIR Butterworth \textbf{ low pass filter} of \nth{5} order with cutoff frequency 10 kHz at -3 dB.

Before the low pass filter usage, the peak at 20 kHz with sideband was present as an unwanted artifact. It could not have been reliably recorded due to the linear frequency response of the sensor up to 10 kHz. At the same time, such frequency is outside the range of any feasible MEMS accelerometer.

\begin{figure}[ht]
    \centering
    \begin{subfigure}[b]{0.44\textwidth}
        \includegraphics[width=\textwidth]{assets/design/Mafaulda-A-time-waveform.png}
        \caption{Temporal domain waveforms}
        \label{fig:design:fault-temporal-waveform}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.55\textwidth}
        \includegraphics[width=\textwidth]{assets/design/Mafaulda-A-spectrum-Y-axis.png}
        \caption{Spectra in radial direction}
        \label{fig:design:fault-spectral-waveform}
    \end{subfigure} 
    \caption{Signals for each fault category with the highest fault severity at 2500 rpm}
\end{figure}

Temporal domain waveforms of the 300 ms signal slice are shown in the graphs in Figure \ref{fig:design:fault-temporal-waveform}. Subplots for radial, tangential, and axial directions are laid out in columns from left to right. Amplitudes vary with limits from $\pm 3\; \mathrm{m/s}^2$ in baseline and misalignment time series up to $\pm 11\;\mathrm{m/s}^2$ in case of severe bearing faults.

The frequency spectrum in Figure \ref{fig:design:fault-spectral-waveform} is obtained by FFT and Hann window of length $2^{14}$. The signal chunk represents an uncertainty box with a duration of approximately 328 ms and a spectral resolution of little over 3 Hz. The graph has been cropped in both axes to make the most important peaks visible.

\subsection{Statistical tests}
%TODO boxplot in featue selection
%TODO images of machines
%TODO

\subsection{Feature extraction}
%TODO - activity diagram, implementation and value counts
